Matching the Time of Emergence of Transboundary Fish Stocks to Lead Time for

### 2021-05-26 Scripts/Initial_Analysis.Rmd
Cleaned code to represent paper estimates

  # Transboundary Index
  Trans_index <- Centroids_id %>% 
    left_join(historical,
              by = c("taxon_key","eez_name","ensemble","neighbour_id")
    ) %>%
    # Re arrenging table for mutate
    gather("variable","value",distance:hd_temp_sd) %>%
    ungroup() %>% 
    mutate(value = ifelse(is.na(value),0,value)) %>% 
    mutate(
      names = ifelse(eez_name ==Unique_EEZs[1],paste("Country_A",variable,sep="_"),paste("Country_B",variable,sep="_")
    )
    ) %>%
    select(-eez_name,-variable) %>%
    spread(names,value) %>% 
    # mutate index
    mutate(
      trans_index = (Country_A_distance/Country_A_hd_temp_sd - Country_B_distance/Country_B_hd_temp_sd)^2
    # )
    ) %>% 
    # Ten years average of results
    group_by(taxon_key,ensemble,neighbour_id) %>% 
    mutate(RMean = rollmean(x = trans_index, 
                            10, 
                            align = "right", 
                            fill = trans_index,
                            na.rm=T)
    ) %>% 
    # filter(year > 1960) %>%
    select(taxon_key,neighbour_id,year,ensemble,trans_index,RMean)

@@

    # Valid ensembles
  # Only perform analysis on ensembles containing all years between 1951 and 2040
    n_years <- mean_sd_index %>% 
      group_by(n_ens) %>% 
      summarise(
        n_year = n()
      ) %>% 
      filter(
        n_year == 149
      )

@@

  # Estimate noice from ensemble members
  historic_sd <- Trans_index %>% 
    filter(year <= Hist_y) %>% 
    # Alternative method sensitivity analysis
    group_by(taxon_key,neighbour_id,year) %>% #Average historic data per year first
    # group_by(taxon_key,neighbour_id,ensemble) %>% #Average historic data per year first
    summarise_at(vars("trans_index"),
                 c(trend= mean),
                 na.rm=T
                 ) %>% 
    group_by(taxon_key,neighbour_id) %>% #Capture ensemble variation
    summarise_at(vars("trend"),
                 c(hist_mean=mean,
                   noise=sd,
                   n_ens = ~n())
                 )

@@

    Raw_Result <- mean_sd_index %>%
      filter(year > Hist_y) %>%
      mutate(
        tresh_one = ifelse(trend > (historic_sd$hist_mean+historic_sd$noise) | trend < (historic_sd$hist_mean-historic_sd$noise),"One",NA),
        tresh_two = ifelse(trend > (historic_sd$hist_mean+(2*historic_sd$noise)) | trend < (historic_sd$hist_mean-(2*historic_sd$noise)),"Two",NA)
      ) %>%
      gather("tresh","emerged",tresh_one,tresh_two)

    taxon <- unique(Raw_Result$taxon_key)
    name_raw <- paste(my_path("R",extra_path = "/Emergence_alt_sens_raw/"),taxon,"_emergence_raw.csv",sep="")

    write_csv(Raw_Result,
              name_raw)

    ToE_Result <- Raw_Result %>% 
    # Method the last year when the index overshoots and never goes back
      filter(is.na(emerged)) %>% 
      group_by(taxon_key, neighbour_id,n_ens,tresh) %>% 
      summarize(emerging_yr = max(year)+1) %>% # last year of NA + 1 makes it the first year of overshooting
      left_join(EEZ_ids,
                by = "neighbour_id") %>%
      ungroup() %>%
      select(taxon_key,eez_name,eez_neighbour,n_ens,emerging_yr,tresh)

  return(ToE_Result)
  }
}

@@

    name <- paste(my_path("R",extra_path = "/Emergence_alt_sens/"),taxon_key,"_emergence.txt",sep="")

>>>>

  # Transboundary Index
  Trans_index <- Centroids_id %>% 
    left_join(historical,
              by = c("taxon_key","eez_name","ensemble","neighbour_id")
    ) %>%
    # Re arrenging table for mutate
    gather("variable","value",distance:hd_temp_sd) %>%
    ungroup() %>% 
    mutate(value = ifelse(is.na(value),0,value)) %>% 
    mutate(
      names = ifelse(eez_name ==Unique_EEZs[1],paste("Country_A",variable,sep="_"),paste("Country_B",variable,sep="_")
    )
    ) %>%
    select(-eez_name,-variable) %>%
    spread(names,value) %>% 
    # mutate index
    mutate(
      trans_index = (Country_A_distance/Country_A_hd_temp_sd - Country_B_distance/Country_B_hd_temp_sd)^2
    # )
    ) %>% 
    # Ten years average of results
    group_by(taxon_key,ensemble,neighbour_id) %>% 
    mutate(RMean = rollmean(x = trans_index, 
                            10, 
                            align = "right", 
                            fill = NA,
                            na.rm=T)
    ) %>% 
    # filter(year > 1960) %>% # drop first 10 years of running mean
    select(taxon_key,neighbour_id,year,ensemble,trans_index,RMean)

@@

    # Valid ensembles
  # Only perform analysis on ensembles containing all years between 1961 and 2040
    n_years <- mean_sd_index %>% 
      group_by(n_ens) %>% 
      summarise(
        n_year = n()
      ) %>% 
      filter(
        n_year == 149
      )

@@

  # Estimate noice from ensemble members
  historic_sd <- Trans_index %>% 
    filter(year <= Hist_y) %>% 
    # Alternative method sensitivity analysis
    # group_by(taxon_key,neighbour_id,year) %>% #s.d. historic data per year first
    group_by(taxon_key,neighbour_id,ensemble) %>% #Average historic data per year first
    summarise_at(vars("RMean"),
                 c(trend = mean),
                 na.rm=T
                 ) %>% 
    group_by(taxon_key,neighbour_id) %>% #Capture ensemble variation
    summarise_at(vars("trend"),
                 c(hist_mean=mean,
                   noise=sd,
                   n_ens = ~n())
                 )

@@

Final_Result <- mean_sd_index %>%
  filter(year > Hist_y) %>%
  mutate(
    tresh_one = ifelse(trend > (historic_sd$hist_mean+historic_sd$noise) | trend < (historic_sd$hist_mean-historic_sd$noise),"One",NA),
    tresh_two = ifelse(trend > (historic_sd$hist_mean+(2*historic_sd$noise)) | trend < (historic_sd$hist_mean-(2*historic_sd$noise)),"Two",NA)
  ) %>%
  gather("tresh","YesNo",tresh_one,tresh_two) %>% 
  filter(!is.na(YesNo)) %>% 
  group_by(taxon_key,neighbour_id,n_ens,tresh) %>%
  summarise(emerging_yr= min(year)) %>%
  # summarise(emerging_yr= max(year)) %>% # Last year that treshold is crossed
  left_join(EEZ_ids,
            by = "neighbour_id") %>%
  ungroup() %>%
  select(taxon_key,eez_name,eez_neighbour,n_ens,emerging_yr,tresh)

  return(Final_Result)
  }
}

@@

    name <- paste(my_path("R",extra_path = "/Emergence_95/"),taxon_key,"_emergence.txt",sep="")
